---
title: "Creation of Phenotypes Summary file"
author: "Yan Holtz"
date: "October 2016"
output:
  rmdformats::readthedown:
    highlight: kate
  html_document:
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document: default
---






[//]: =========================================================================================================================


#1/ Introduction
This file is a supplementary data attached with the publication.  
It describes the analysis of the phenotypic data.
These data are organized experiment per experiment (distinct files). There are 7 experiments, in different places and year.

This RMD file allows to:  
- charge every file  
- compute blup of each individuals  
- compute heritability for each trait  
- merge all experiments together  
- put explicit names to variable  
--> A summary file is made


We need a few libraries
```{r results='hide', message=FALSE, warning=FALSE}
library(xtable)
library(gdata)
library(lme4)
library(RColorBrewer)
my_colors=brewer.pal(8, "Set2") 
```

Let's upload every file
```{r}
# Watch out, to reproduct analysis, you have to update the path.
my_path="/Users/yan/Dropbox/Publi_Fusariose/ANALYSIS_REPRO/DATA/PHENOTYPE"
setwd(my_path)
# Cappelles
CAP13=read.xls("Cappelle_2013_données élaborées.xlsx", sheet = 5, header = TRUE, na.strings="NA") 
# Grisolles
GRI11=read.xls("Grisolles_2011_données élaborées.xlsx", sheet = 1, header = TRUE, na.strings="NA") 
GRI13=read.xls("Grisolles_2013_données élaborées.xlsx", sheet = 3, header = TRUE) 
GRI15=read.xls("Grisolles_2015_données élaborées.xlsx", sheet = 3, header = TRUE) 
#Lectoure
LEC14=read.xls("Lectoure_2014_données élaborées.xlsx", sheet = 3, header = TRUE) 
# Montbequi
BEQ11=read.xls("Monbequi_2011_données élaborées.xlsx", sheet = 2, header = TRUE) 
# Montbartier
BAR14=read.xls("Montbartier_2014_données élaborées.xlsx", sheet = 2, header = TRUE) 
# DON
CAP13_DON=read.xls("DON données élaborées.xlsx", sheet = 1, header = TRUE) 
GRI13_DON=read.xls("DON données élaborées.xlsx", sheet = 2, header = TRUE) 
LEC14_DON=read.xls("DON données élaborées.xlsx", sheet = 3, header = TRUE) 
# All individuals that exist for this pop:
my_indiv=read.table("all_dic2*Silur.txt", header=F)[,1]
my_indiv=c("dic2","silur",as.character(my_indiv) )
```

[//]: =========================================================================================================================








[//]: =========================================================================================================================

#2/ Useful Functions

A function to calculate heritability:
```{r}
calculate_herit=function(my_col, my_data){
  
  # Combien de rep2?
  r2=nrow(my_data[ which(my_data$rep==2 & !is.na(my_data[,my_col])), ])

  # si 0, alors je retourne NA pour l'hérit:
  if(r2==0){ return("NA") ; continue }
  
  #Calcul hérit
  aov<- lmer ( my_data[,my_col] ~ (1|geno)  , data=my_data)
  VL<-as.numeric(VarCorr(aov)$geno)
  VRes<-as.numeric(attr(VarCorr(aov),"sc"))^2
  hdeux=round(VL/(VL + VRes) , 6)
  return(hdeux)
}
```

Same function, but with inoculation date in the model:
```{r}
calculate_herit_ino=function(my_col, my_data, my_ino){
  
  # Combien de rep2?
  r2=nrow(my_data[ which(my_data$rep==2 & !is.na(my_data[,my_col])), ])

  # si 0, alors je retourne NA pour l'hérit:
  if(r2==0){ return("NA") ; continue }
  
  #Calcul hérit
  aov<- lmer ( my_data[,my_col] ~ (1|geno) + my_data[,my_ino]  , data=my_data)
  VL<-as.numeric(VarCorr(aov)$geno)
  VRes<-as.numeric(attr(VarCorr(aov),"sc"))^2
  hdeux=round(VL/(VL + VRes) , 6)
  return(hdeux)
}
```

A function that return BLUPs of individuals. Takes no cofactor
```{r}
calculate_blup_nocofactor=function(my_col, my_data){
  aov<- lmer ( my_data[,my_col] ~ (1|geno)   , data=my_data)
  blup <- ranef(aov, condVar = TRUE)
  tmp=data.frame(rownames(blup$geno), blup$geno[,1])
  colnames(tmp)[2]=paste( colnames(my_data)[my_col], "blup", sep="_")
  return(tmp)
  }
```

A function that return BLUPs of individuals. It **DOES** take innoculation date into account
```{r}
calculate_blup=function(my_col, my_ino, my_data){
  aov<- lmer ( my_data[,my_col] ~ (1|geno) + my_data[,my_ino]  , data=my_data)
  blup <- ranef(aov, condVar = TRUE)
  tmp=data.frame(rownames(blup$geno), blup$geno[,1])
  colnames(tmp)[2]=paste( colnames(my_data)[my_col], "blup", sep="_")
  return(tmp)
  }
```

A function that return BLUPs of individuals. It **DOES** take **innoculation** date AND **height** into account
```{r}
calculate_blup_height=function(my_col, my_ino, my_height, my_data){
  aov<- lmer ( my_data[,my_col] ~ (1|geno) + my_data[,my_ino] + my_data[,my_height]  , data=my_data)
  blup <- ranef(aov, condVar = TRUE)
  tmp=data.frame(rownames(blup$geno), blup$geno[,1])
  colnames(tmp)[2]=paste( colnames(my_data)[my_col], "blup", sep="_")
  return(tmp)
  }
```

A function that return BLUPs of individuals. It **DOES** take  **height** into account **only**
```{r}
calculate_blup_heightONLY=function(my_col,  my_height, my_data){
  aov<- lmer ( my_data[,my_col] ~ (1|geno) + my_data[,my_height]  , data=my_data)
  blup <- ranef(aov, condVar = TRUE)
  tmp=data.frame(rownames(blup$geno), blup$geno[,1])
  colnames(tmp)[2]=paste( colnames(my_data)[my_col], "blup", sep="_")
  return(tmp)
  }
```

[//]: =========================================================================================================================










[//]: =========================================================================================================================


#3/ Cappelle 2013
Give better column names to the dataset & clean data
```{r}
# Rename columns:
my_names=c("geno", "rep", "CAP13_EPI", "CAP13_FLO", "CAP13_HEI", "CAP13_VER", "CAP13_INO", "CAP13_INO2", "CAP13_PEPI300", "CAP13_PEPI400", "CAP13_PEPI550", "CAP13_PEPIAUDPC", "CAP13_NEPIL300", "CAP13_NEPIL400", "CAP13_NEPIL550", "CAP13_NEPILAUDPC", "CAP13_PEPIL300", "CAP13_PEPIL400", "CAP13_PEPIL550", "CAP13_PEPILAUDPC")
colnames(CAP13)=my_names
CAP13$geno=gsub("TT06DC ", "TT06DC.", CAP13$geno)
# Delete the rep2 of PEPI400, NEPIL400 et PEPIL400 which are wrong
CAP13$CAP13_PEPI400[CAP13$rep=="2"]=NA
CAP13$CAP13_NEPIL400[CAP13$rep=="2"]=NA
CAP13$CAP13_PEPIL400[CAP13$rep=="2"]=NA
```


Let's calculate and print heritabilities (with and without inoc date effect):
```{r}
# Calcul héritabilité:
herit_CAP13=matrix(0,2,ncol(CAP13)-2)
colnames(herit_CAP13)=my_names[3:length(my_names)]
num=0
for(i in c(3:ncol(CAP13))){
  num=num+1
  a=calculate_herit(i, CAP13)
  b=calculate_herit_ino(i, CAP13, 7)
  herit_CAP13[ ,num]=c(a,b)
}
herit_CAP13=apply(herit_CAP13 , 2 , as.numeric)
```

```{r my_table1, results='asis' , echo=FALSE}
print(xtable(herit_CAP13), type = "html", include.rownames = F , comment=FALSE)
```

Change the format of the data: distincts columns for rep1 and rep2 (one line per genotype)
```{r}
a=CAP13[CAP13$rep==1 , c(-2)]
colnames(a)[2:ncol(a)]=paste( my_names[3:length(my_names)], "rep1", sep="_" )
b=CAP13[CAP13$rep==2 , c(-2)]
colnames(b)[2:ncol(b)]=paste( my_names[3:length(my_names)], "rep2", sep="_" )
CAP13_clean=merge(a, b, by.x=1, by.y=1 , all=T)
CAP13_clean=CAP13_clean[ , -which(colnames(CAP13_clean)%in%c("CAP13_PEPI400_rep2","CAP13_NEPIL400_rep2","CAP13_PEPIL400_rep2"))]
```

Then we have to calculate the BLUP of every genotype for every variable and add it to the "clean" table :
```{r}
for(i in c(3:ncol(CAP13))){
  
  # Cas particulier pour  les 400
  if( i %in% which(colnames(CAP13)%in%c("CAP13_PEPI400","CAP13_NEPIL400","CAP13_PEPIL400"))){ 
    tmp=aggregate( CAP13[,i], by=list(CAP13[,1]) , mean, na.rm=T) 
    colnames(tmp)[2]=paste(colnames(CAP13)[i],"_blup",sep="")
  }else{
    tmp=calculate_blup_height(i, 7, 5, CAP13)
  }
  
  CAP13_clean=merge(CAP13_clean , tmp, by.x=1, by.y=1, all.x=T)
  }
```



[//]: =========================================================================================================================












[//]: =========================================================================================================================


#4/ Grisolles 2011
Give better column names to the dataset
```{r}
# Rename columns:
my_names=c("geno", "rep", "GRI11_HEI", "GRI11_EPI", "GRI11_NOT300", "GRI11_NOT350", "GRI11_NOT450", "GRI11_NOT500", "GRI11_NOT550", "GRI11_NOTAUDPC", "GRI11_NEPIL300", "GRI11_NEPIL350", "GRI11_NEPIL450", "GRI11_NEPIL500", "GRI11_NEPIL550", "GRI11_NEPILAUDPC", "GRI11_NANT" )
colnames(GRI11)=my_names
GRI11$geno=gsub("TT06DC ", "TT06DC.", GRI11$geno)
GRI11$geno=gsub("DIC2", "dic2", GRI11$geno)
GRI11$geno=gsub("Silur", "silur", GRI11$geno)
```


Let's calculate and print heritabilities. No innoculation date provided so heritability is calculated only once.
```{r}
# Calcul héritabilité:
herit_GRI11=matrix(0,1,ncol(GRI11)-2)
colnames(herit_GRI11)=my_names[3:length(my_names)]
num=0
for(i in c(3:ncol(GRI11))){
  num=num+1
  a=calculate_herit(i, GRI11) 
  herit_GRI11[1,num]=a
  }
```

```{r my_table2, results='asis' , echo=FALSE}
print(xtable(herit_GRI11), type = "html", include.rownames = F , comment=FALSE)
```

To finish, we change the format of the data: distincts columns for rep1 and rep2
```{r}
a=GRI11[GRI11$rep==1 , c(-2)]
colnames(a)[2:ncol(a)]=paste( my_names[3:length(my_names)], "rep1", sep="_" )
b=GRI11[GRI11$rep==2 , c(-2)]
colnames(b)[2:ncol(b)]=paste( my_names[3:length(my_names)], "rep2", sep="_" )
GRI11_clean=merge(a, b, by.x=1, by.y=1 , all=T)
```

Then we have to calculate the BLUP of every genotype for every variable and add it to the "clean" table.
Since there is no height provided here, we duplicate the height of rep1
```{r}
# Duplicate height of rep1.
aaa=aggregate(GRI11$GRI11_HEI , by=list(GRI11$geno) , mean, na.rm=T)
GRI11=merge(GRI11, aaa, by.x=1 , by.y=1, all.x=T)
GRI11$GRI11_HEI=GRI11$x
GRI11=GRI11[ , -ncol(GRI11)]
# Calcultate blup
for(i in c(4: (ncol(GRI11)-1) )){
  tmp=calculate_blup_heightONLY(i, 3, GRI11)
  GRI11_clean=merge(GRI11_clean , tmp, by.x=1, by.y=1, all.x=T)
  }
```
[//]: =========================================================================================================================












[//]: =========================================================================================================================


#5/ Grisolles 2013
Give better column names to the dataset
```{r}
# Rename columns:
GRI13=GRI13[,c(-2,-12)]
my_names=c("geno", "rep", "GRI13_EPI", "GRI13_FLO", "GRI13_HEI", "GRI13_INO", "GRI13_PEPI350", "GRI13_PEPI450", "GRI13_PEPI550", "GRI13_PEPIAUDPC", "GRI13_NEPIL350", "GRI13_NEPIL450", "GRI13_NEPIL550", "GRI13_NEPILAUDPC", "GRI13_PEPIL350","GRI13_PEPIL450","GRI13_PEPIL550","GRI13_PEPILAUDPC",   "GRI13_NOT350", "GRI13_NOT450", "GRI13_NOT550", "GRI13_NOTAUDPC")
colnames(GRI13)=my_names
GRI13$geno=gsub("TT06DC ", "TT06DC.", GRI13$geno)
```


Let's calculate and print heritabilities:
```{r}
# Calcul héritabilité:
herit_GRI13=matrix(0,2,ncol(GRI13)-2)
colnames(herit_GRI13)=my_names[3:length(my_names)]
num=0
for(i in c(3:ncol(GRI13))){
  num=num+1
  a=calculate_herit(i, GRI13) 
  b=calculate_herit_ino(i, GRI13, 6)
  herit_GRI13[ ,num]=c(a,b)
  }
```

```{r my_table3, results='asis' , echo=FALSE}
print(xtable(herit_GRI13), type = "html", include.rownames = F , comment=FALSE)
```

To finish, we change the format of the data: distincts columns for rep1 and rep2
```{r}
a=GRI13[GRI13$rep==1 , c(-2)]
colnames(a)[2:ncol(a)]=paste( my_names[3:length(my_names)], "rep1", sep="_" )
b=GRI13[GRI13$rep==2 , c(-2)]
colnames(b)[2:ncol(b)]=paste( my_names[3:length(my_names)], "rep2", sep="_" )
GRI13_clean=merge(a, b, by.x=1, by.y=1 , all=T)
```

Then we have to calculate the BLUP of every genotype for every variable and add it to the "clean" table :
```{r}
for(i in c(3:ncol(GRI13))){
  tmp=calculate_blup_height(i, 6, 5, GRI13)
  GRI13_clean=merge(GRI13_clean , tmp, by.x=1, by.y=1, all.x=T)
  }
```

[//]: =========================================================================================================================
















[//]: =========================================================================================================================


#6/ Grisolles 2015
Give better column names to the dataset
```{r}
# Rename columns:
my_names=c("geno", "rep", "GRI15_EPI", "GRI15_HEI", "GRI15_INO", "GRI15_NOT350", "GRI15_NOT550", "GRI15_NOTAUDPC", "GRI15_PEPI350", "GRI15_PEPI550", "GRI15_PEPIAUDPC", "GRI15_NEPIL350", "GRI15_NEPIL550", "GRI15_NEPILAUDPC", "GRI15_PEPIL350", "GRI15_PEPIL550", "GRI15_PEPILAUDPC" )
colnames(GRI15)=my_names
GRI15$geno=gsub("TT06DC ", "TT06DC.", GRI15$geno)
GRI15$geno=gsub(" L3", "", GRI15$geno)
GRI15$geno=gsub(" L2", "", GRI15$geno)

```

Let's calculate and print heritabilities:
```{r}
# Calcul héritabilité:
herit_GRI15=matrix(0,2,ncol(GRI15)-3)
colnames(herit_GRI15)=my_names[c(3,4,6:length(my_names))]
num=0
for(i in c(3,4,6:length(my_names))){
  num=num+1
  a=calculate_herit(i, GRI15) 
  b=calculate_herit_ino(i, GRI15, 5)
  herit_GRI15[ ,num]=c(a,b)
  }
```


```{r my_table4, results='asis' , echo=FALSE}
print(xtable(herit_GRI15), type = "html", include.rownames = F , comment=FALSE)
```

To finish, we change the format of the data: distincts columns for rep1 and rep2
```{r}
a=GRI15[GRI15$rep==1 , c(-2)]
colnames(a)[2:ncol(a)]=paste( my_names[3:length(my_names)], "rep1", sep="_" )
b=GRI15[GRI15$rep==2 , c(-2)]
colnames(b)[2:ncol(b)]=paste( my_names[3:length(my_names)], "rep2", sep="_" )
GRI15_clean=merge(a, b, by.x=1, by.y=1 , all=T)
```

Then we have to calculate the BLUP of every genotype for every variable and add it to the "clean" table :
```{r}
for(i in c(3,4,6:length(my_names)) ){
  tmp=calculate_blup_height(i, 5, 4, GRI15)
  GRI15_clean=merge(GRI15_clean , tmp, by.x=1, by.y=1, all.x=T)
  }
```

[//]: =========================================================================================================================











[//]: =========================================================================================================================


#7/ Lectoure 2014
Give better column names to the dataset
```{r}
# Rename columns:
my_names=c("geno", "rep", "LEC14_EPI", "LEC14_HEI", "LEC14_EXT", "LEC14_INO", "LEC14_PEPI300", "LEC14_PEPI400", "LEC14_PEPI550", "LEC14_PEPIAUDPC", "LEC14_NEPIL300", "LEC14_NEPIL400", "LEC14_NEPIL550", "LEC14_NEPILAUDPC", "LEC14_PEPIL300", "LEC14_PEPIL400", "LEC14_PEPIL550", "LEC14_PEPILAUDPC" )
colnames(LEC14)=my_names
LEC14$geno=gsub("TT06DC ", "TT06DC.", LEC14$geno)
LEC14$geno=gsub(" L3", "", LEC14$geno)
LEC14$geno=gsub(" L2", "", LEC14$geno)
LEC14$geno=gsub("Silur", "silur", LEC14$geno)
# Delete weird individuals
LEC14=LEC14[-which(LEC14$geno=="TT06DC.40.45" | LEC14$geno=="TT06DC.42.12" | LEC14$geno=="TT06DC.42.14" | LEC14$geno=="TT06DC.44.62" | LEC14$geno=="TT06DC.44.64")  , ]

```


Let's calculate and print heritabilities:
```{r}
# Calcul héritabilité:
herit_LEC14=matrix(0,2,ncol(LEC14)-3)
colnames(herit_LEC14)=my_names[c(3,5:length(my_names))]
num=0
for(i in c(3,5:ncol(LEC14))){
  num=num+1
  a=calculate_herit(i, LEC14) 
  b=calculate_herit_ino(i, LEC14, 6)
  herit_LEC14[ ,num]=c(a,b)
  }
```


```{r my_table5, results='asis' , echo=FALSE}
print(xtable(herit_LEC14), type = "html", include.rownames = F , comment=FALSE)
```

To finish, we change the format of the data: distincts columns for rep1 and rep2
```{r}
a=LEC14[LEC14$rep==1 , c(-2)]
colnames(a)[2:ncol(a)]=paste( my_names[3:length(my_names)], "rep1", sep="_" )
b=LEC14[LEC14$rep==2 , c(-2)]
colnames(b)[2:ncol(b)]=paste( my_names[3:length(my_names)], "rep2", sep="_" )
LEC14_clean=merge(a, b, by.x=1, by.y=1 , all=T)
```

Then we have to calculate the BLUP of every genotype for every variable and add it to the "clean" table :
Oui duplicate the height since there is only one rep.

```{r}
# Duplicate height of rep1.
aaa=aggregate(LEC14$LEC14_HEI , by=list(LEC14$geno) , mean, na.rm=T)
LEC14=merge(LEC14, aaa, by.x=1 , by.y=1, all.x=T)
LEC14$LEC14_HEI=LEC14$x
LEC14=LEC14[ , -ncol(LEC14)]
# Calculate blup
for(i in c(3,5:length(my_names))  ){
  print(i)
  tmp=calculate_blup_height(i, 6, 4,  LEC14)
  LEC14_clean=merge(LEC14_clean , tmp, by.x=1, by.y=1, all.x=T)
  }
```

[//]: =========================================================================================================================











[//]: =========================================================================================================================

#8/ Montbequi 2011

Give better column names to the dataset
```{r}
# Rename columns:
my_names=c("geno", "rep", "BEQ11_FLO",  "BEQ11_HEI", "BEQ11_LCE", "BEQ11_COM", "BEQ11_PEPI350", "BEQ11_PEPI450", "BEQ11_PEPIAUDPC", "BEQ11_NOT300", "BEQ11_NOT350", "BEQ11_NOT450", "BEQ11_NOT500", "BEQ11_NOTAUDPC", "BEQ11_NEPIL300", "BEQ11_NEPIL350", "BEQ11_NEPIL450", "BEQ11_NEPIL500", "BEQ11_NEPILAUDPC" )
colnames(BEQ11)=my_names
BEQ11$geno=gsub("TT06DC ", "TT06DC.", BEQ11$geno)

```

No heritability since there is *NO repetition*
No innoculation date effect since it is not provided
To finish, we change the format of the data: just aggregate data for dic2 and silur. Other are all right since there is no rep.
Il faudrait faire un modele pour prendre en compte la hauteur quand même!
```{r}
#Pas propre
BEQ11=aggregate( cbind(BEQ11_FLO, BEQ11_HEI, BEQ11_LCE, BEQ11_COM, BEQ11_PEPI350, BEQ11_PEPI450, BEQ11_PEPIAUDPC, BEQ11_NOT300, BEQ11_NOT350, BEQ11_NOT450, BEQ11_NOT500, BEQ11_NOTAUDPC, BEQ11_NEPIL300, BEQ11_NEPIL350, BEQ11_NEPIL450, BEQ11_NEPIL500, BEQ11_NEPILAUDPC) ~ geno, data=BEQ11, mean)
BEQ11_clean=BEQ11
```

[//]: =========================================================================================================================










[//]: =========================================================================================================================


#9/ Montbartier 2014

Give better column names to the dataset
```{r}
# Rename columns:
my_names=c("geno", "rep", "BAR14_EPI",  "BAR14_ROUY", "BAR14_NOT600", "BAR14_INO", "BAR14_INO2", "BAR14_INO3", "BAR14_NEPIL", "BAR14_EXT", "BAR14_PEPI350", "BAR14_PEPI400", "BAR14_PEPI550", "BAR14_PEPI600", "BAR14_PEPIAUDPC", "BAR14_NEPIL300", "BAR14_NEPIL400", "BAR14_NEPIL550", "BAR14_NEPIL600", "BAR14_NEPILAUDPC", "BAR14_PEPIL300", "BAR14_PEPIL400", "BAR14_PEPIL550", "BAR14_PEPIL600", "BAR14_PEPILAUDPC")
colnames(BAR14)=my_names
BAR14$geno=gsub("TT06DC ", "TT06DC.", BAR14$geno)
BAR14$geno=gsub(" L3", "", BAR14$geno)
BAR14$geno=gsub(" L2", "", BAR14$geno)
BAR14$geno=gsub("SILUR", "silur", BAR14$geno)
BAR14$geno=gsub("Dic 2", "dic2", BAR14$geno)
```

Let's calculate and print heritabilities:
```{r}
# Calcul héritabilité:
herit_BAR14=matrix(0,2,ncol(BAR14)-5)
colnames(herit_BAR14)=my_names[c(3,4,5,9:length(my_names))]
num=0
for(i in c(3,4,5,9:ncol(BAR14))){
  num=num+1
  a=calculate_herit(i, BAR14) 
  b=calculate_herit_ino(i, BAR14, 6)
  herit_BAR14[ ,num]=c(a,b)
  }
```


```{r my_table1115, results='asis' , echo=FALSE}
print(xtable(herit_BAR14), type = "html", include.rownames = F , comment=FALSE)
```


To finish, we change the format of the data: distincts columns for rep1 and rep2
```{r}
a=BAR14[BAR14$rep==1 , c(-2)]
colnames(a)[2:ncol(a)]=paste( my_names[3:length(my_names)], "rep1", sep="_" )
b=BAR14[BAR14$rep==2 , c(-2)]
colnames(b)[2:ncol(b)]=paste( my_names[3:length(my_names)], "rep2", sep="_" )
BAR14_clean=merge(a, b, by.x=1, by.y=1 , all=T)
```

Then we have to calculate the BLUP of every genotype for every variable and add it to the "clean" table.

Since the heigt is not provided, I can put the height of CAP13.

```{r}
# Add average height of CAP13.
my_height=aggregate(CAP13$CAP13_HEI , by=list(CAP13$geno), mean)
colnames(my_height)=c("geno","height")
BAR14=merge(BAR14, my_height, by.x=1, by.y=1 , all.x=T)

# Calculate blups
for(i in c(3,4,5,9:length(my_names))  ){
  tmp=calculate_blup_height(i, 6, 26, BAR14)
  BAR14_clean=merge(BAR14_clean , tmp, by.x=1, by.y=1, all.x=T)
  }
```


[//]: =========================================================================================================================









[//]: =========================================================================================================================


#10/ DON

## Cappelle 2013
```{r}
# Clean data
# geno "TT06DC.38.34" hand-made changed to rep1 and rep2
CAP13_DON=CAP13_DON[,3:5]
colnames(CAP13_DON)=c("geno", "rep", "CAP13_DON")
CAP13_DON$rep[CAP13_DON$rep=="RepM"]="Rep1"
CAP13_DON$geno=gsub("TT06DC ", "TT06DC.", CAP13_DON$geno)
CAP13_DON$geno=gsub("Silur", "silur", CAP13_DON$geno)
CAP13_DON$rep=gsub("Rep", "", CAP13_DON$rep)
# On enleve 2 individus douteux: 
CAP13_DON=CAP13_DON[-which(CAP13_DON$geno=="TT06DC.44.24" | CAP13_DON$geno=="TT06DC.42.36") ,]


# Heritability
a=calculate_herit(3, CAP13_DON) 
herit_CAP13=cbind(herit_CAP13,data.frame(CAP13_DON=c(a,"NA")))

# each rep in a column
a=CAP13_DON[CAP13_DON$rep==1 , c(-2)]
b=CAP13_DON[CAP13_DON$rep==2 , c(-2)]
CAP13_DON_clean=merge(a, b, by.x=1, by.y=1 , all=T)
colnames(CAP13_DON_clean)=c("geno","CAP13_DON_rep1", "CAP13_DON_rep2")

# add blups
# I did a first try using the height in the model. It was quite disapoiting. So let's try taking into acount the % of spikelet with fusariose.
CAP13_DON=merge(CAP13_DON, CAP13[,c("geno","rep","CAP13_PEPIL550")], by.x=c(1,2), by.y=c(1,2) , all.x=T)
# calculate blup
aov<- lmer ( CAP13_DON ~ (1|geno) + CAP13_PEPIL550 , data=CAP13_DON)
blup <- ranef(aov, condVar = TRUE)
tmp=data.frame(geno=rownames(blup$geno), CAP13_DON_blup=blup$geno[,1])
CAP13_DON_clean=merge(CAP13_DON_clean, tmp, by.x=1, by.y=1, all.x=T)
```



## Grisolles 2013
```{r}
# Clean data
GRI13_DON=GRI13_DON[,3:5]
colnames(GRI13_DON)=c("geno", "rep", "GRI13_DON")
GRI13_DON$rep[GRI13_DON$rep=="RepM"]="Rep1"
GRI13_DON$geno=gsub("TT06DC ", "TT06DC.", GRI13_DON$geno)
GRI13_DON$geno=gsub("Silur", "silur", GRI13_DON$geno)
GRI13_DON$rep=gsub("Rep", "", GRI13_DON$rep)
# On enleve une donnée indiqué comme louche
GRI13_DON=GRI13_DON[-which(GRI13_DON$geno=="TT06DC.38.53" & GRI13_DON$GRI13_DON==706),]
# On enleve 4 individus pourris
GRI13_DON=GRI13_DON[-which(GRI13_DON$geno=="TT06DC.39.21" | GRI13_DON$geno=="TT06DC.39.24" | GRI13_DON$geno=="TT06DC.38.50" | GRI13_DON$geno=="TT06DC.42.44") ,]


# Heritability
a=calculate_herit(3, GRI13_DON) 
herit_GRI13=cbind(herit_GRI13,data.frame(GRI13_DON=c(a,"NA")))

# each rep in a column
a=GRI13_DON[GRI13_DON$rep==1 , c(-2)]
b=GRI13_DON[GRI13_DON$rep==2 , c(-2)]
GRI13_DON_clean=merge(a, b, by.x=1, by.y=1 , all=T)
colnames(GRI13_DON_clean)=c("geno","GRI13_DON_rep1", "GRI13_DON_rep2")

# add blups
# We use the PEPIL 550 to better estimate the blups?
GRI13_DON=merge(GRI13_DON, GRI13[ , c("geno","rep","GRI13_PEPIL550")], by.x=c(1,2), by.y=c(1,2) , all.x=T)
# calculate blup
aov<- lmer ( GRI13_DON ~ (1|geno) +  GRI13_PEPIL550 , data=GRI13_DON)
blup <- ranef(aov, condVar = TRUE)
tmp=data.frame(geno=rownames(blup$geno), GRI13_DON_blup=blup$geno[,1])
GRI13_DON_clean=merge(GRI13_DON_clean, tmp, by.x=1, by.y=1, all.x=T)
```


## Lectoure 2014
Il faut juste moyenner les valeurs pour Dic2 et Silur.
Il n'y a pas de rep concernant DON pour cette expé, meme si il y avait des rep au champ!
```{r}
# Clean data
LEC14_DON=aggregate(LEC14_DON$DON, by=list(geno=as.factor(LEC14_DON$geno)) , mean)
colnames(LEC14_DON)=c("geno", "LEC14_DON")
LEC14_DON$geno=gsub("TT06DC ", "TT06DC.", LEC14_DON$geno)
LEC14_DON$geno=gsub("Dic2", "dic2", LEC14_DON$geno)
LEC14_DON$geno=gsub("Silur", "silur", LEC14_DON$geno)
LEC14_DON_clean=LEC14_DON

```

[//]: =========================================================================================================================







[//]: =========================================================================================================================


#11/ Heritability
Let's show the heritability of every trait of this study. Several barplots are presented. From top to bottom: Cappelle, Grisolles 2011, 2013 and 2015, Lectoure, Montbartier. From left to right: Agronomical features, Resistance of type 1, resistance of type 2.

## Graphic representation
```{r, fig.height=9}
# Start barplot
par(mfrow=c(6,3))
par(mar=c(1,2,1,0))

# Cappelles 2013
dat=as.data.frame(herit_CAP13)
colnames(dat)=gsub("CAP13_","", colnames(dat))
a=as.matrix( dat[1,c("EPI","FLO","HEI","VER")] )
barplot(a, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1), ylab="CAPPELLE 2013" )
b=as.matrix( dat[1,c("PEPI300","PEPI400","PEPI550","PEPIAUDPC")] )
barplot(b, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) )
c=as.matrix( dat[1,c("NEPIL300","NEPIL400","NEPIL550","NEPILAUDPC","PEPIL300","PEPIL400","PEPIL550","PEPILAUDPC")] )
barplot(c, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) )

# Grisolles 2011
dat=as.data.frame(herit_GRI11)
colnames(dat)=gsub("GRI11_","", colnames(dat))
a=as.matrix( dat[1,c("EPI","HEI")] )
barplot(a, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1), ylab="GRISOLLES 2011" )
b=as.matrix( dat[1,c("NOT300","NOT350","NOT450","NOT500","NOT550","NOTAUDPC")] )
barplot(b, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) )
c=as.matrix( dat[1,c("NEPIL300","NEPIL350","NEPIL450","NEPIL500","NEPIL550","NEPILAUDPC")] )
barplot(c, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) )

# Grisolles 2013
dat=as.data.frame(herit_GRI13)
colnames(dat)=gsub("GRI13_","", colnames(dat))
a=as.matrix( dat[1,c("EPI","FLO","HEI")] )
barplot(a, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1), ylab="GRISOLLES 2013" )
b=as.matrix( dat[1,c("PEPI350","PEPI450","PEPI550","PEPIAUDPC","NOT350", "NOT450", "NOT550", "NOTAUDPC")] )
barplot(b, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) )
c=as.matrix( dat[1,c("NEPIL350","NEPIL450","NEPIL550","NEPILAUDPC","PEPIL350","PEPIL450","PEPIL550","PEPILAUDPC")] )
barplot(c, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) )

# Grisolles 2015
dat=as.data.frame(herit_GRI15)
colnames(dat)=gsub("GRI15_","", colnames(dat))
a=as.matrix( dat[1,c("EPI","HEI")] )
barplot(a, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) , ylab="GRISOLLES 2015")
b=as.matrix( dat[1,c("NOT350","NOT550","NOTAUDPC","PEPI350","PEPI550", "PEPIAUDPC")] )
barplot(b, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) )
c=as.matrix( dat[1,c("NEPIL350","NEPIL550","NEPILAUDPC","PEPIL350","PEPIL550","PEPILAUDPC")] )
barplot(c, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) )

# Lectoure 2014
dat=as.data.frame(herit_LEC14)
colnames(dat)=gsub("LEC14_","", colnames(dat))
a=as.matrix( dat[1,c("EPI","EXT")] )
barplot(a, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) , ylab="LECTOURE 2014")
b=as.matrix( dat[1,c("PEPI300","PEPI400","PEPI550","PEPIAUDPC")] )
barplot(b, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) )
c=as.matrix( dat[1,c("NEPIL300","NEPIL400", "NEPIL550","NEPILAUDPC","PEPIL300","PEPIL400","PEPIL550","PEPILAUDPC")] )
barplot(c, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) )

# Pas de rep pour Montbequi

# Montbartier 14
dat=as.data.frame(herit_BAR14)
colnames(dat)=gsub("BAR14_","", colnames(dat))
a=as.matrix( dat[1,c("EPI","ROUY","EXT")] )
barplot(a, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) , ylab="LECTOURE 2014")
b=as.matrix( dat[1,c("NOT600", "PEPI350","PEPI400","PEPI550","PEPI600","PEPIAUDPC")] )
barplot(b, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) )
c=as.matrix( dat[1,c("NEPIL300","NEPIL400", "NEPIL550","NEPIL600","NEPILAUDPC","PEPIL300","PEPIL400","PEPIL600","PEPILAUDPC")] )
barplot(c, xlab="", col="grey", cex.names=0.7, las=2, ylim=c(0,1) )

```



## Rough values
*Cappelles 2013*
```{r my_table6, results='asis' , echo=FALSE}
print(xtable(herit_CAP13), type = "html", include.rownames = F , comment=FALSE)
```

*Grisolles 2011*
```{r my_table7, results='asis' , echo=FALSE}
print(xtable(herit_GRI11), type = "html", include.rownames = F , comment=FALSE)
```

*Grisolles 2013*
```{r my_table8, results='asis' , echo=FALSE}
print(xtable(herit_GRI13), type = "html", include.rownames = F , comment=FALSE)
```

*Grisolles 2015*
```{r my_tablell8, results='asis' , echo=FALSE}
print(xtable(herit_GRI15), type = "html", include.rownames = F , comment=FALSE)
```

*Lectoure 2014*
```{r my_table9, results='asis' , echo=FALSE}
print(xtable(herit_LEC14), type = "html", include.rownames = F , comment=FALSE)
```

*Montbequi 2011*
Pas de rep?

*Montbartier 2014*
```{r my_table10, results='asis' , echo=FALSE}
print(xtable(herit_BAR14), type = "html", include.rownames = F , comment=FALSE)
```


## Effect of Innoculation date
Using innoculation date always gives lower heritability... 
```{r}
# Show
par(mfrow=c(2,3), mar=c(8,2,2,2))
my_list=list(herit_CAP13[,c(1:18)], herit_GRI13[,-21] , herit_GRI15, herit_LEC14, herit_BAR14)
lapply(my_list, function(x){ barplot(as.matrix(x),beside=T, las=2, col=my_colors[c(1,2)] )}  )
```

## Conclusion
- Flowering time & Height have really good heritability --> we can trust our experimental design.--
- **BUT** Grisolles 2011: low heritability for every trait --> no trust --> We have to remove this field for further analysis.--
- AUDPC and last notation are always the best option.--
- Heritability are medium ~0.5.--
- Taking into account the innoculation date has low effect on te heritability, except fro Cappelle 2013. (Except for flowering data, what is logical)
- DON has a low heritability (~0.25 only...)

[//]: =========================================================================================================================







[//]: =========================================================================================================================
#12/ Output phenotype table
Now we need to merge all these data together for further study (QTL)

```{r}

# Concatenate every file
FINAL=my_indiv
for(i in list(CAP13_clean, CAP13_DON_clean, GRI11_clean, GRI13_clean, GRI13_DON_clean, GRI15_clean, LEC14_clean, LEC14_DON_clean, BEQ11_clean, BAR14_clean)){
  #print(head(colnames(i)))
  FINAL=merge(FINAL, i, by.x=1, by.y=1, all.x=T)
   }
colnames(FINAL)[1]="geno"

# I remove the variable concerning innoculation date:
FINAL=FINAL[ , -grep( "INO", colnames(FINAL) ) ]

write.table(FINAL, file="/Users/yan/Dropbox/Publi_Fusariose/ANALYSIS_REPRO/DATA/PHENOTYPE/phenotypage_all_fusa.csv", quote=F, row.names=F, col.names=T, sep=";")
```

Nombre de variable final??
```{r}
dim(FINAL)
```
J'ai bien les 0 dans les noms?
```{r}
#FINAL$geno
```

Duplicated genotyped?
```{r}
table(FINAL$geno)[table(FINAL$geno)>1]
```
[//]: =========================================================================================================================






[//]: =========================================================================================================================

#13/ Feature of experiments
Let's summarize the feature of each experiment in a table:

```{r}
# start table
bilan=data.frame( matrix(0,8,7) )
colnames(bilan)=c("CAP13", "GRI11", "GRI13", "GRI15", "LEC14", "BEQ11","BAR14")
rownames(bilan)=c("# obs", "# rep", "# geno", "DON?", "PEPI?", "NOT?", "NEPIL?", "PEPIL?")
all=list(CAP13, GRI11, GRI13, GRI15, LEC14, BEQ11, BAR14)
all_clean=list(CAP13_clean, GRI11_clean, GRI13_clean, GRI15_clean, LEC14_clean, BEQ11_clean, BAR14_clean)

# Number of obs per exp?
bilan[1,]=unlist(lapply(all, nrow ))
# Number of ind rep:
bilan[2,]=unlist(lapply(all, function(x){nrow(x[x$rep==2, ])} ))
# Number of genotypes
bilan[3,]=unlist(lapply(all_clean, nrow) )
# Don available?
bilan[4,]=c("Yes","-","Yes","-","Yes","-","-")
# PEPI avail
bilan[5,]=c("Yes","-","Yes","Yes","Yes","Yes","Yes")
# NOT avail
bilan[6,]=c("-","Yes","Yes","Yes","-","Yes","Yes")
# NEPIL
bilan[7,]=c("Yes","Yes","Yes","Yes","Yes","Yes","Yes")
# PEPIL
bilan[8,]=c("Yes","-","Yes","Yes","Yes","-","Yes")
```

```{r my_table111ee5, results='asis' , echo=FALSE}
print(xtable(bilan), type = "html", include.rownames = T , comment=TRUE)
```

[//]: =========================================================================================================================












[//]: =========================================================================================================================

#14/ Effect of innoculation date?

##Cappelle 2013
Pour Cappelle, on observe un léger effet (plus de maladie lorsque l'Inoc à lieu tot). Mais cette effet n'est pas signif.
```{r}
# show
par(mfrow=c(1,3), mar=c(5,4,2,2))
boxplot( CAP13$CAP13_PEPIAUDPC ~ CAP13$CAP13_INO, ylab="PEPI (AUDPC)", xlab="Innoculation date", col=my_colors)
boxplot( CAP13$CAP13_PEPILAUDPC ~ CAP13$CAP13_INO, ylab="PEPIL (AUDPC)", xlab="Innoculation date", col=my_colors)
boxplot( CAP13$CAP13_NEPILAUDPC ~ CAP13$CAP13_INO, ylab="NEPIL (AUDPC)", xlab="Innoculation date", col=my_colors)
# Check if the innoculation date has an effect on the disease observation. I return the pvalues of the anova test for the innoculation effect. If < 0.05, we have a significant effect.
for(i in c(9:ncol(CAP13))){
  model=lm( CAP13[,i] ~ geno + CAP13_INO , data=CAP13)
  a=anova(model)[2,5]
  model2=lm( CAP13[,i] ~ geno + CAP13_INO2 , data=CAP13)
  a=anova(model)[2,5]
  b=anova(model2)[2,5]
  print(paste(colnames(CAP13)[i], a, b,sep="  -  "))
}

```

## Grisolles 2011
Pas de données des dates d'Inoc

## Grisolles 2013
One more time, samples that are innoculated earlier show more symptoms. The difference is significant.
```{r}
# show
par(mfrow=c(1,4), mar=c(5,4,2,2))
boxplot( GRI13$GRI13_PEPIAUDPC ~ GRI13$GRI13_INO, ylab="PEPI (AUDPC)", xlab="Innoculation date", col=my_colors)
boxplot( GRI13$GRI13_PEPILAUDPC ~ GRI13$GRI13_INO, ylab="PEPIL (AUDPC)", xlab="Innoculation date", col=my_colors)
boxplot( GRI13$GRI13_NEPILAUDPC ~ GRI13$GRI13_INO, ylab="NEPIL (AUDPC)", xlab="Innoculation date", col=my_colors)
boxplot( GRI13$GRI13_NOTAUDPC ~ GRI13$GRI13_INO, ylab="NOT (AUDPC)", xlab="Innoculation date", col=my_colors)
# Check if the innoculation date has an effect on the disease observation. I return the pvalues of the anova test for the innoculation effect. If < 0.05, we have a significant effect.
for(i in c(8:ncol(GRI13))){
  model=lm( GRI13[,i] ~ geno + GRI13_INO , data=GRI13)
  a=anova(model)[2,5]
  print(paste(colnames(GRI13)[i],a, sep=" : "))
}
```

## Grisolles 2015
Hard to interpret. There is a significant difference between innoculation date. But it is not ordered. May be the date 1-->6 are not in the good order?

```{r}
# show
par(mfrow=c(1,3), mar=c(5,4,2,2))
boxplot( GRI15$GRI15_PEPILAUDPC ~ GRI15$GRI15_INO, ylab="PEPIL (AUDPC)", xlab="Innoculation date", col=my_colors)
boxplot( GRI15$GRI15_NEPILAUDPC ~ GRI15$GRI15_INO, ylab="NEPIL (AUDPC)", xlab="Innoculation date", col=my_colors)
boxplot( GRI15$GRI15_NOTAUDPC ~ GRI15$GRI15_INO, ylab="NOT (AUDPC)", xlab="Innoculation date", col=my_colors)
# Significant?
for(i in c(5:ncol(GRI15))){
  model=lm( GRI15[,i] ~ geno + GRI15_INO , data=GRI15)
  a=anova(model)[2,5]
  print(paste(colnames(GRI15)[i],a, sep=" : "))
}
```

## Lectoure 2014
Visually, a strong effect as usual (later inoc=less disease). But not significant...

```{r}
# show
par(mfrow=c(1,3), mar=c(5,4,2,2))
boxplot( LEC14$LEC14_PEPILAUDPC ~ LEC14$LEC14_INO, ylab="PEPIL (AUDPC)", xlab="Innoculation date", col=my_colors)
boxplot( LEC14$LEC14_NEPILAUDPC ~ LEC14$LEC14_INO, ylab="NEPIL (AUDPC)", xlab="Innoculation date", col=my_colors)
boxplot( LEC14$LEC14_PEPIAUDPC ~ LEC14$LEC14_INO, ylab="PEPI (AUDPC)", xlab="Innoculation date", col=my_colors)
# Significant?
for(i in c(7:ncol(LEC14))){
  model=lm( LEC14[,i] ~ geno + LEC14_INO , data=LEC14)
  a=anova(model)[2,5]
  print(paste(colnames(LEC14)[i],a, sep=" : "))
}
```

## Montbequi 2011
No innoculation date provided.

## Montbartier 2014

```{r}
# show
par(mfrow=c(1,3), mar=c(5,4,2,2))
boxplot( BAR14$BAR14_PEPILAUDPC ~ BAR14$BAR14_INO, ylab="PEPIL (AUDPC)", xlab="Innoculation date", col=my_colors)
boxplot( BAR14$BAR14_NEPILAUDPC ~ BAR14$BAR14_INO, ylab="NEPIL (AUDPC)", xlab="Innoculation date", col=my_colors)
boxplot( BAR14$BAR14_PEPILAUDPC ~ BAR14$BAR14_INO, ylab="PEPIL (AUDPC)", xlab="Innoculation date", col=my_colors)
# Significant?
for(i in c(11:ncol(BAR14))){
  model=lm( BAR14[,i] ~ geno + as.factor(BAR14_INO) , data=BAR14)
  a=anova(model)[2,5]
  print(paste(colnames(BAR14)[i],a, sep=" : "))
}
```


[//]: =========================================================================================================================








[//]: =========================================================================================================================

#15/ Effect of height in blup?

On calcule les blups avec et sans effet de la hauteur:
```{r}
NOCOF=calculate_blup_nocofactor(11, CAP13)
SANS=calculate_blup(11, 7, CAP13)
AVEC=calculate_blup_height(11, 7, 5, CAP13)
```

Distribution de la variable et des blups :
```{r}
par(mfrow=c(1,3))
hist(CAP13[,11] , col=rgb(0.1,0.8,0.1,0.4) , xlab="Original value of CAP13_PEPI550" , breaks=20, main="")
hist(SANS[,2] , col=rgb(0.3,0.8,0.6,0.4) , xlab="Blup without height" , breaks=20, main="")
hist(AVEC[,2] , col=rgb(0.1,0.4,0.6,0.4) , xlab="Blup with height" , breaks=20, main="")
```

On voit que la variable de base n'est pas du tout normale. Lorsque l'on passe au BLUP, la variable devient quasi-normale et répartie de -20 à 20 ou -40 à 40, ce qui est beaucoup plus classique.

Plot de la relation entre note de base (rep1) et blup?
```{r}
AA=merge(CAP13[ which(CAP13$rep==1) ,c(1,11)]  , SANS, by.x=1, by.y=1, all.x=T  )
AA=merge(AA, AVEC, by.x=1, by.y=1, all.x=T)
par(mfrow=c(1,2))
plot(AA[ ,2] , AA[,3] , pch=20 , col=rgb(0.3,0.2,0.6,0.4) , xlab="Original value of CAP13_PEPI550", ylab="Blup without height" , cex=1.3)
plot(AA[ ,2] , AA[,4] , pch=20 , col=rgb(0.3,0.2,0.6,0.4) , xlab="Original value of CAP13_PEPI550", ylab="Blup without height" , cex=1.3)
```


Plot de la relation entre les 2:
```{r}
par(mfrow=c(1,1))
plot(SANS[,2] , AVEC[,2] , pch=20 , col=rgb(0.3,0.2,0.6,0.4) , xlab="Blup without height", ylab="Blup without height" , cex=1.3)
cor(SANS[,2] , AVEC[,2])
```
On a une très bonne corrélation entre les 2 blups, ce qui est plutôt rassurant. Reste à voir si les variations nous paraissent logiques.

On rappelle que les plantes **hautes** sont censées être plus **résistante**.
Essayons de trouver une plante très grande et une très petite, avec la même note de PEPI550.

Le génotype 42.26 est très petit, avec des hauteurs de 65 et 75 pour rep1 et 2 resp. Concernant sont PEPI550, il est de 100 pour la rep1, NA pour la rep2.
Le génotype 39.26 est beaucoup plus grand (1m25 et 1m). Il a aussi un PEPI550 de 100 pour la rep1 uniquement.

Sans prendre en compte de cofacteur, ces 2 génotypes devraient avoir le même blup, puisqu'ils ont les meme scores de PEPI550. On peut le vérifier:

```{r}
NOCOF[NOCOF[,1]=="TT06DC.42.26" | NOCOF[,1]=="TT06DC.39.26" ,]
```

Ces 2 génotypes ont donc un score équivalent. Pourtant le 42.26 est petit, donc a plus de chance d'être touché par la Fusa. On pourrait donc dire qu'il est probablement plus résistant que le 39.26. On s'attend donc à ce qu'il ait un blup un peu plus bas! Vérifions le!

```{r}
AVEC[AVEC[,1]=="TT06DC.42.26" | AVEC[,1]=="TT06DC.39.26" ,]
```

On voit que c'est clairement le cas (5.42 vs -1.7). Donc le BLUP travaille correctement et devrait nous permettre d'améliorer nos QTLs!

[//]: =========================================================================================================================







[//]: =========================================================================================================================

#15/ Effect of PEPIL in DON blups?

on va prendre GRI13 comme exemple.

Quelle est la relation entre le % d'épillets touchés et la valeur de don?  
--> quasiment pas de corrélation. Du coup ça risque de ne pas trop changer les blups j'imagine.
```{r}
plot(CAP13_DON$CAP13_PEPIL550 , CAP13_DON$CAP13_DON, xlab="PEPIL 550" , ylab="DON" , pch=20 , col=rgb(0.6,0.1,0.0,0.4), cex=1.3)
abline(lm(CAP13_DON$CAP13_DON ~ CAP13_DON$CAP13_PEPIL550) , col="grey")
```


Je calcule les blups avec et sans prise en compte du PEPIL

```{r}
# calculate blup without 
aov<- lmer ( GRI13_DON ~ (1|geno)  , data=GRI13_DON)
blup <- ranef(aov, condVar = TRUE)
without=data.frame(geno=rownames(blup$geno), GRI13_DON_blup=blup$geno[,1])

# calculate blup WITH
aov<- lmer ( GRI13_DON ~ (1|geno) +  GRI13_PEPIL550 , data=GRI13_DON)
blup <- ranef(aov, condVar = TRUE)
with=data.frame(geno=rownames(blup$geno), GRI13_DON_blup=blup$geno[,1])

# merge all
AA=merge( CAP13_DON_clean , without , by.x=1 , by.y=1 , all.x=T)
AA=merge( AA, with , by.x=1 , by.y=1 , all.x=T)
```

Relation entre les 2 blups? --> C'est très corrélé!
```{r}
# plot
plot(AA[,4] , AA[,5], xlab="blup basic", ylab="blup taking PEPIL into account" , pch=20, col=rgb(0.9,0.3,0.1,0.3) )
cor(AA[,4] , AA[,5], use="complete.obs")
```


[//]: =========================================================================================================================



