---
title: "Polyphénol - part 1: créer un tableau bilan"
author: "Yan Holtz"
date: "October 2016"
output:
  rmdformats::readthedown:
    highlight: kate
  html_document:
    toc: yes
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document: default
---






[//]: =========================================================================================================================


#1/ Introduction
This file is a supplementary data attached with the publication.  
The only goal is to create a summary file that gives the quantity of phenol compounds observed for each genotype. 
The data are organized in 2 years (2013 and 2014)

This RMD file allows to:
- charge every file
- compute blup of each individuals
- compute heritability for each trait
- merge all experiments together

We need a few libraries
```{r results='hide', message=FALSE, warning=FALSE}
library(xtable)
library(gdata)
library(lme4)
library(RColorBrewer)
library(plotly)
library(FactoMineR)
my_colors=brewer.pal(8, "Set2") 
```

Let's upload every file
```{r}
# Watch out, to reproduct analysis, you have to update the path.
#my_path="../../DATA/PHENOTYPE"
#setwd(my_path)
# Metabolomique
DATA13=read.xls("../../DATA/PHENOTYPE/Metabolomique_Data.xlsx",  sheet = 1, header = TRUE) 
DATA14=read.xls("../../DATA/PHENOTYPE/Metabolomique_Data.xlsx",  sheet = 2, header = TRUE) 
# All individuals that exist for this pop:
my_indiv=read.table("../../DATA/PHENOTYPE/all_dic2_Silur.txt", header=F)[,1]

```
[//]: =========================================================================================================================








[//]: =========================================================================================================================

#2/ Useful Functions

A function to calculate heritability:
```{r}
calculate_herit=function(my_col, my_data){
  
  aov<- lmer ( my_data[,my_col] ~ (1|geno)  , data=my_data)
  VL<-as.numeric(VarCorr(aov)$geno)
  VRes<-as.numeric(attr(VarCorr(aov),"sc"))^2
  hdeux=round(VL/(VL + VRes) , 6)
  return(hdeux)
}
```


A function that return BLUPs of individuals. 
```{r}
calculate_blup=function(my_col, my_data){
  aov<- lmer ( my_data[,my_col] ~ (1|geno) , data=my_data)
  blup <- ranef(aov, condVar = TRUE)
  tmp=data.frame(rownames(blup$geno), blup$geno[,1])
  colnames(tmp)[2]=paste( colnames(my_data)[my_col], "blup", sep="_")
  return(tmp)
  }
```

[//]: =========================================================================================================================










[//]: =========================================================================================================================


#3/ Year 2013
Clean the dataset
```{r}
# Rename columns:
DATA13$geno=as.factor(gsub("TT06DC", "TT06DC.", DATA13$geno))
```

How many different individuals in this files? (counting dic2 and silur)
```{r}
nlevels(DATA13$geno)
```

Always 3 reps / genotypes? --> Yes indeed !
```{r}
table(DATA13$geno)[table(DATA13$geno)!=3]
```

How many variables?
```{r}
ncol(DATA13)-2
```


Let's calculate and print heritabilities (with and without inoc date effect):
```{r}
# Calcul héritabilité:
herit_DATA13=matrix(0,1,ncol(DATA13)-2)
colnames(herit_DATA13)=colnames(DATA13)[3:length(colnames(DATA13))]
num=0
for(i in c(3:ncol(DATA13))){
  num=num+1
  a=calculate_herit(i, DATA13)
  herit_DATA13[1 ,num]=c(a)
  }
```

Display these heritabilities in a table:
```{r my_tabletrhtr1, results='asis' , echo=FALSE}
print(xtable(herit_DATA13), type = "html", include.rownames = F , comment=FALSE)
```

And show a borplot with these heritabilities:
```{r}
plot_ly(x=colnames(herit_DATA13), y=herit_DATA13[1,], type="bar")%>%
        layout(
          yaxis = list(title = 'Heritability'),
          xaxis = list(tickfont = list(size=2))
        )
```

Heritabilities are incredibly high. That means repetitions are strongly correlated one each other. So we'are going to compute the blups of genotypes and use only these blups for the QTLs.


Now we calculate the BLUP of every genotype for every variable and add it to a summary table in which we will add the 2 years:
```{r}
FINAL=my_indiv
for(i in c(3:ncol(DATA13))){
  tmp=calculate_blup(i,  DATA13)
  FINAL=merge(FINAL , tmp, by.x=1, by.y=1, all.x=T)
}
colnames(FINAL)=gsub("blup","blup13", colnames(FINAL))
```

[//]: =========================================================================================================================











[//]: =========================================================================================================================


#4/ Year 2014
Clean the dataset
```{r}
# Rename columns:
DATA14$geno=gsub("Dic2", "dic2", DATA14$geno)
DATA14$geno=gsub("Silur", "silur", DATA14$geno)
DATA14$geno=gsub(" L3", "", DATA14$geno)
DATA14$geno=as.factor(gsub("TT06DC ", "TT06DC.", DATA14$geno))
DATA14=DATA14[ , -3]
```

How many different individuals in this files? (counting dic2 and silur)
```{r}
nlevels(DATA14$geno)
```

Always 3 reps / genotypes? --> Yes indeed !
```{r}
table(DATA14$geno)[table(DATA14$geno)!=3]
```

How many variables?
```{r}
ncol(DATA14)-2
```

Let's calculate and print heritabilities (with and without inoc date effect):
```{r}
# Calcul héritabilité:
herit_DATA14=matrix(0,1,ncol(DATA14)-2)
colnames(herit_DATA14)=colnames(DATA14)[3:length(colnames(DATA14))]
num=0
for(i in c(3:ncol(DATA14))){
  num=num+1
  a=calculate_herit(i, DATA14)
  herit_DATA14[1 ,num]=c(a)
  }
```

Display these heritabilities in a table:
```{r my_tabledfef, results='asis' , echo=FALSE}
print(xtable(herit_DATA14), type = "html", include.rownames = F , comment=FALSE)
```

And show a borplot with these heritabilities:
```{r}
plot_ly(x=colnames(herit_DATA14), y=herit_DATA14[1,], type="bar")%>%
        layout(
          yaxis = list(title = 'Heritability'),
          xaxis = list(tickfont = list(size=2))
        )
```

Heritabilities are incredibly high. That means repetitions are strongly correlated one each other. So we'are going to compute the blups of genotypes and use only these blups for the QTLs.


Now we calculate the BLUP of every genotype for every variable and add it to a summary table in which we will add the 2 years:
```{r}
for(i in c(3:ncol(DATA14))){
  tmp=calculate_blup(i,  DATA14)
  FINAL=merge(FINAL , tmp, by.x=1, by.y=1, all.x=T)
}
colnames(FINAL)=gsub("blup13","yoyoyo", colnames(FINAL))
colnames(FINAL)=gsub("blup","blup14", colnames(FINAL))
colnames(FINAL)=gsub("yoyoyo","blup13", colnames(FINAL))
colnames(FINAL)[1]="geno"
```

Let's write this table for the qTL step.
```{r}
write.table(FINAL, file="../../DATA/PHENOTYPE/phenotypage_all_metabolomique.csv", quote=F, row.names=F, col.names=T, sep=";")
```

Nombre de variable final??
```{r}
dim(FINAL)
```
J'ai bien les 0 dans les noms?
```{r}
FINAL$geno
```

Duplicated genotyped?
```{r}
table(FINAL$geno)[table(FINAL$geno)>1]
```


[//]: =========================================================================================================================










